---
title: "final crime project"
author: "YI-Hung, Lu, Hawa Toumbou, Hatice Erdogan"
date: "Due: Doc.14 2023"
output: 
  word_document: 
    toc: yes
    toc_depth: 2
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Research Question:

How did the COVID-19 pandemic specifically influence the trends of property and violent crimes in Washington, D.C., and what are the underlying factors contributing to these changes?

Null Hypothesis:

The COVID- 19 pandemic significantly influenced the trends of property and violent crimes in Washington, D.C., and any observed changes in these crime rates are not distinctively attributable to the pandemic. 

Alternative Hypothesis:

The COVID-19 pandemic significantly influenced the trends of property and violent crimes in Washington, D.C., with identifiable changes in the patterns and rates of these crimes attributable to pandemic-related social, economic, and policy factors.

Crime Patterns in Urban Centers During the COVID-19 Pandemic: A Case Study of Washington D.C.

Abstract:

This study examines the multifaceted socio-economic impacts of the COVID-19 pandemic, with a specific focus on how these changes have influenced crime patterns in urban environments, particularly Washington D.C. The paper integrates theoretical perspectives from urban sociology and criminology, emphasizing the intersection between pandemic-induced socioeconomic shifts and urban crime dynamics.

Introduction:

The COVID-19 pandemic has caused a major shift in the world and is strongly affecting many aspects of human life and society. The pandemic was so dire that it caused governments to take drastic measures in order to try and control the exposure as well as spread of the disease. Many countries adopted various lockdown measures, such as limit outside or public activity, closures of business, isolation in homes with family and further quarantine within your home if infected. 

It ignited one of the largest global economic crises unparalleled in recent history. We see the effects of this in the dramatic increase in inequality within nations throughout the world. Lower income households, of course, felt the strain moreso which will cause them to need more time to recover. The economic strain forced many to deplete their savings for survival if they were lucky enough to have savings in the first place. In April 2020 alone, the job market saw a staggering loss of over 20 million jobs, with the closure of businesses across sectors further exacerbating the situation.

	During and after the pandemic there has been a change in social dynamics. We have seen unprecedented socioeconomic changes globally, with profound impacts on urban centers. Lockdowns and social distancing measures significantly altered routine activities and social interactions. These changes could affect crime patterns, as theories like the routine activity theory suggest that crime is a function of daily activities and interactions. Social and economic stressors have definitely been an effect of the lockdown and pandemic. Economic hardships were at the forefront of the pandemic which can exacerbate factors that lead to certain crimes. Increases in unemployment and financial stress could lead to a rise in crimes driven by economic desperation.  

The psychological impact of prolonged isolation and stress was also significant. Extended periods of stress and anxiety, typical in such crises, can lead to increased irritability and aggression, potentially contributing to a rise in domestic violence and mental health-related incidents. While some types of crimes may have decreased in frequency, others may have found new ground or increased due to the changing circumstances. In urban cities like Washington D.C., a multifaceted array of factors influences crime rates. These include socioeconomic disadvantages, population density, educational opportunities, family structure, unemployment, access to weapons, and migration and demographic changes.  

Washington D.C is one of the most influential cities in the nation. Its high concentration of government institutions, universities, and business scene has long attracted out-of-towners wanting to relocate to a more lively environment rich with culture. Washington D.C. has been going through migration and demographic changes. During the heightened period of the pandemic, the population influx slowed but soon after the lockdown protocols were lifted, the shift changed increasing gentrification within the nation's capital and the surrounding counties. 

Prior to the pandemic, movement in and out of the nation's capital was normal due to the universities and job opportunities. The city was prone to seasonal fluctuations, with populations taking a slight dip during summer due to universities closing. But due to job openings and internship season, there was still a big number of people moving in during the months of May and June. Throughout the years there has been an overall upwards trend in the population change. There was a huge dip between the months of February and May of 2020. Stringent lockdown measures were implemented between February and May of 2020, leading to a significant dip in the movement of people. The bustling streets quieted, and the usual influx of professionals and interns was noticeably reduced. The pandemic's impact was felt across all aspects of city life, from the economy to the healthcare system, which faced unprecedented challenges during these months. During that period, people moved back home from school. Some people decided to move closer to their family given that they were able to with the eventual development of telework. 

The COVID-19 pandemic significantly influenced the transition to digital platforms and telework, with profound effects on employment, particularly in lower-income communities. During the pandemic, the United States experienced substantial job losses, and the work-experience unemployment rate in 2020 was approximately 15.0%, similar to the rate during the Great Recession in 2010. This impact however was not felt across the all income tiers. Disparities in digital literacy and technological knowledge further marginalized lower-income workers, making it challenging for them to adapt to remote working environments and leverage digital tools effectively. The pandemic-induced shift to digital and telework environments illuminated and intensified the digital divide. 

Background Data on Unemployment Rates:
2018: 5.6% year average 
2019: 4.9% year average
March 2020: 6.10% (beginning of the pandemic impact)
2020 (Pandemic Onset):
April 2020: 10.40%, (sharp rise as pandemic impacts begin)
Subsequent months in 2020 saw high unemployment rates, peaking in July at 9.40%, and gradually decreasing to 7.00% by December.

2021 (Pandemic):
The unemployment rate remained high but gradually decreased throughout 2021, starting at 6.60% in January and fluctuating before decreasing to 6.40% by October.

2022 (Post-Pandemic Recovery):
The year 2022 showed a noticeable recovery with a decreasing trend in unemployment rates, starting from 6.30% in January, reducing to 4.20% in April, and fluctuating around 4%-5% towards the end of the year.
2023:
In 2023, the unemployment rate continued to fluctuate slightly but remained around 4%-5%, indicating stabilization post-pandemic​

These figures reflect the significant economic impact of the COVID-19 pandemic, with a notable spike in unemployment rates during the initial months of the pandemic, followed by a gradual recovery over the subsequent years. The gradual decrease seen in December of 2020 shows the impact of telework and transitioning more into the digital age. 
Gentrification is defined as the influx of higher income population into a geographic area while lower-income, generally minority residents are forced out due to increasing cost of living. U.S. inflation rate for 2022 was 8.00%, a 3.3% increase from 2021. U.S. inflation rate for 2021 was 4.70%, a 3.46% increase from 2020. U.S. inflation rate for 2020 was 1.23%, a 0.58% decline from 2019. U.S. inflation rate for 2019 was 1.81%, a 0.63% decline from 2018.The cost of living in the Washington D.C. area has increased tremendously since the outbreak and continues to rise. 
The overall Consumer Price Index (CPI) rose by 3.3% in 2023 which tells us that the upward trend continues to climb. A major contributor is food prices, which have increased notably, with a 4.2% rise over the year. This can impact lower-income households more severely, possibly leading to increased financial pressure. The rising cost of living, especially in essentials like food, shelter, and education, can exacerbate economic inequalities and financial hardships. This, in turn, could potentially lead to an increase in crime as individuals and families struggle to cope with the financial pressures.

Unlike most American cities, Washington D.C is experiencing significant displacement of low-income residents due to gentrification, as reported by the Institute of Metropolitan Opportunity. Ward 6 in D.C., including Capitol Hill, Navy Yard, and the Southwest Waterfront, shows dramatic demographic changes. Low-income populations in parts of these neighborhoods have decreased significantly (up to 75% in some areas). Areas like Shaw and Columbia Heights have seen a considerable drop in low-income populations (up to 57% in Shaw) and a substantial increase in the white population. The government and local officials want to attract more business and investment opportunities for the city but without the people of the city and that is how this becomes harmful. 

Opportunity Zones (OZs) is a federal tax incentive, and the benefits are applied to federal capital gains. It is controlled by the federal government to receive financial incentives for investments. These zones, intended to boost economic development in distressed neighborhoods, are often selected by state governors and local politicians from eligible census tracts. This can be harmful in several ways, particularly in the context of its impact on black communities. 
Displacement of Low-Income Residents are one of the primary concerns with the implementation of  OZs. The influx of investment aimed at redevelopment and commercialization can lead to an increase in property values and living costs, making it unaffordable for the original residents to continue living in these areas. OZs can accelerate the process of gentrification, where wealthier individuals move into a neighborhood, often changing its character and dynamics. This process can lead to a cultural and social displacement of the existing community members, where the needs and preferences of new, often more affluent residents, overshadow those of the longstanding residents. In Washington, D.C., 25 out of 97 low-income communities and 19 contiguous census tracts were designated as OZs. While OZs are intended to stimulate economic growth in distressed areas, it presents a risk of accelerated gentrification and its potential negative impacts, particularly the displacement of lower-income residents. 
The Washington D.C. housing market from 2018 to 2023 has experienced significant changes in terms of home prices, mortgage rates, and their potential impact on the city's dynamics, including the aspect of crime rates. The housing market in Washington D.C. has seen a period of steady increase in home prices. As of 2023, the median home price stands at around $667,932, marking a significant rise over the past several years. This rise is attributed to the city's strong economy and job market, along with high demand for housing in the area. The increase in home prices has led to a situation where homes have been sold at prices significantly higher than the listed price, especially evident in early 2022 where bidding wars were common​. 
Rapid increases in housing costs and mortgage rates can contribute to economic disparities and housing affordability issues, which in some cases, may indirectly influence crime rates. Areas experiencing rapid gentrification and displacement of lower-income residents could see changes in their social dynamics, which might impact crime rates as well. 
This paper aims to analyze these shifts and changes in crime rates in Washington D.C., providing a comprehensive overview of the pandemic's multifaceted impact on the nation's capital.

```{r}
library(tidyverse)
library(stringr)      
library(dplyr)
library(ggplot2)	
library(tidyr)        
library(reshape2)
library(readr)		
library(forcats)
library(car)
```

The dataset are search from the opendata.dc.gov.
Clear data
```{r}
data1 <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_2018.csv", delim = ",", col_types = cols(
  REPORT_DAT = col_character(),
  METHOD = col_character(),
  OFFENSE = col_character(),
  MONTH_YR = col_character(),
  WARD = col_double()
))
data1
data11 <- data1[data1$MONTH_YR != "-", ]
data11
# Assuming your data frame is named 'your_data'
data111 <- data11[, !colnames(data11) %in% "...6"]

data2 <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_2019.csv", delim = ",", col_types = cols(
  REPORT_DAT = col_character(),
  METHOD = col_character(),
  OFFENSE = col_character(),
  MONTH_YR = col_character(),
  WARD = col_double()
))
data3 <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_Incidents_in_2020.csv", delim = ",", col_types = cols(
  REPORT_DAT = col_character(),
  METHOD = col_character(),
  OFFENSE = col_character(),
  MONTH_YR = col_character(),
  WARD = col_double()
))
data4 <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_Incidents_in_2021.csv", delim = ",", col_types = cols(
  REPORT_DAT = col_character(),
  METHOD = col_character(),
  OFFENSE = col_character(),
  MONTH_YR = col_character(),
  WARD = col_double()
))
data5 <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_Incidents_in_2022.csv", delim = ",", col_types = cols(
  REPORT_DAT = col_character(),
  METHOD = col_character(),
  OFFENSE = col_character(),
  MONTH_YR = col_character(),
  WARD = col_double()
))
data6 <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_Incidents_in_2023.csv", delim = ",", col_types = cols(
  REPORT_DAT = col_character(),
  METHOD = col_character(),
  OFFENSE = col_character(),
  MONTH_YR = col_character(),
  WARD = col_double()
))
data22 <- data2[data2$MONTH_YR != "--", ]
data33 <- data3[data3$MONTH_YR != "-", ]
data44 <- data4[data4$MONTH_YR != "-", ]
data55 <- data5[data5$MONTH_YR != "-", ]
data66 <- data6[data6$MONTH_YR != "-", ]
```

Combine the six dataset to be a final dataset.
```{r}
data_final <- rbind(data111, data22,data33, data44,data55,data66)
data_final
write.csv(data_final, "Crime_2018-2023.csv", row.names = FALSE)
```

```{r}
crime <- read_csv("/Users/pikaqiu/Desktop/AU /class /R studio/Crime_2018-2023.csv")
crime
```

```{r}
# Separate into month and year 
crime <- crime %>%
  mutate(YEAR = year(as.Date(REPORT_DAT, format = "%Y/%m/%d")),
         MONTH = month(as.Date(REPORT_DAT, format = "%Y/%m/%d")))
```

```{r}
crime$YEAR <- factor(crime$YEAR, levels = c("2018", "2019", "2020", "2021", "2022", "2023"))
```

```{r}
head(crime)
```

```{r}
total_crimes <- crime %>%
  group_by(YEAR) %>%
  summarize(total_crimes = n())
total_crimes
```
```{r}
ggplot(total_crimes, aes(x = YEAR, y = total_crimes, group = 1)) +
  geom_line(color = "blue") +
  geom_point(color = "blue", size = 2) +
  geom_text(aes(label = scales::comma(`total_crimes`), vjust = -0.5)) +
  labs(title = "Total Crime by Year",
       x = "Year",
       y = "Total Crimes") +
  theme_minimal()
```
We see a very sharp decrease in 2020 from almost 34k to almost 28k, then increase in 2021 to 28,303 and then decrease again in 2022 to 27136 then a very sharp increase in 2023 to 32,213, almost reaching the level before the pandemic.

```{r}
percentage_change_2022_to_2023 <- crime %>%
  group_by(YEAR) %>%
  summarize(total_crimes = n()) %>%
  filter(YEAR %in% c("2022", "2023")) %>%
  summarise(percentage_change = ((total_crimes[2] - total_crimes[1]) / total_crimes[1]) * 100)

# Calculate percentage change
percentage_change_2022_to_2023
```
```{r}
percentage_change_2019_to_2020 <- crime %>%
  group_by(YEAR) %>%
  summarize(total_crimes = n()) %>%
  filter(YEAR %in% c("2019", "2020")) %>%
  summarise(percentage_change = ((total_crimes[2] - total_crimes[1]) / total_crimes[1]) * 100)

# Display the result
print(percentage_change_2019_to_2020)
```


```{r}
offense_count <- crime %>%
  group_by(YEAR, OFFENSE) %>%
  count()
offense_count
```
```{r}
# Filter data for the years 2020 and 2023
offenses_2020 <- crime %>%
  filter(YEAR == "2020") %>%
  group_by(OFFENSE) %>%
  summarize(total_crimes = n())

offenses_2023 <- crime %>%
  filter(YEAR == "2023") %>%
  group_by(OFFENSE) %>%
  summarize(total_crimes = n())

# Create horizontal bar charts for each year
ggplot(offenses_2020, aes(x = total_crimes, y = reorder(OFFENSE, total_crimes), fill = OFFENSE, color = OFFENSE)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::comma(`total_crimes`), hjust = -0.1, vjust = 0.5)) +
  labs(title = "Total Crimes by Offense in 2020",
       x = "Total Crimes",
       y = "Offense") +
  theme_minimal() +
  theme(legend.position = "none")  

```
```{r}
ggplot(offenses_2023, aes(x = total_crimes, y = reorder(OFFENSE, total_crimes), fill = OFFENSE, color = OFFENSE)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::comma(`total_crimes`), hjust = -0.1, vjust = 0.5)) +
  labs(title = "Total Crimes by Offense in 2023",
       x = "Total Crimes",
       y = "Offense") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove legend
```
```{r}
ggplot(offense_count, aes(x = factor(YEAR), y = n, fill = OFFENSE)) +
  geom_bar(stat = "identity") +
  labs(title = "Crime Frequency by Year and Offense",
       x = "Year",
       y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "right") +
  theme(text = element_text(size = 10), element_line(size = 1)) +
  scale_x_discrete(labels = seq(2018, 2023, 1))
```
We see a decrease in theft overall with the beginning of the pandemic and increase in homicide and motor vehicle theft after the pandemic, 2023. 

```{r}
month_count <- crime %>%
  group_by(YEAR, MONTH) %>%
  count(OFFENSE) 
#  summarize(sum(n))
month_count
```

```{r}
options(repr.plot.width = 8, repr.plot.height = 6) 

ggplot(crime, aes(x = YEAR, color = OFFENSE)) +
  geom_line(stat = "count", aes(group = interaction(OFFENSE, 1)), size = 1) +
  geom_point(stat = "count", aes(group = interaction(OFFENSE, 1)), shape = 16, size = 2) +
  labs(title = "Incidents of Crimes Over Years",
       x = "Year",
       y = "Count of Incidents",
       color = "Offense Type") +
  theme(text = element_text(size = 9),element_line(size = 1))
```
Theft/other decreases with the beginning of the pandemic and then gradually increases after the pandemic. Theft/auto increases slightly in 2021 and decreases ever since then. Homicide shows gradual increase during the pandemic and keeps increasing. Motor vehicle theft is also one of the offenses that shows only increase after the pandemic. 

```{r}
theft_from_auto<- crime %>%
  filter(OFFENSE == "THEFT F/AUTO")

# Create a line graph with bold circles at each year
ggplot(theft_from_auto, aes(x = YEAR)) +
  geom_line(stat = "count", aes(group = 1), color = "red") +
  geom_point(stat = "count", aes(group = 1), shape = 16, size = 3, color = "red") +
  labs(title = "Incidents of Theft from Auto Over Years",
       x = "Year",
       y = "Count of Incidents")
```
We see a sharp decline with the beginning of the pandemic, then increase in 2021 but decrease ever since. 

```{r}
theft_other <- crime %>%
  filter(OFFENSE == "THEFT/OTHER")

# Create a line graph with bold circles at each year
ggplot(theft_other, aes(x = YEAR)) +
  geom_line(stat = "count", aes(group = 1), color = "red") +
  geom_point(stat = "count", aes(group = 1), shape = 16, size = 3, color = "red") +
  labs(title = "Incidents of Theft/Other Over Years",
       x = "Year",
       y = "Count of Incidents")
```
Theft shows a gradual decrease from 2019 to 2020, stays steady during the pandemic and then increases in 2022.

```{r}
homicide <- crime %>%
  filter(OFFENSE == "HOMICIDE")

# Create a line graph with bold circles at each year
ggplot(homicide, aes(x = YEAR)) +
  geom_line(stat = "count", aes(group = 1), color = "red") +
  geom_point(stat = "count", aes(group = 1), shape = 16, size = 3, color = "red") +
  labs(title = "Homicide Over Years",
       x = "Year",
       y = "Count of Incidents")
```
Homicide shows an upward trend with the beginning of the pandemic then decreases in 2022 and gradually increases in 2023.

```{r}
mtv_theft <- crime %>%
  filter(OFFENSE == "MOTOR VEHICLE THEFT")

# Create a line graph with bold circles at each year
ggplot(mtv_theft, aes(x = YEAR)) +
  geom_line(stat = "count", aes(group = 1), color = "red") +
  geom_point(stat = "count", aes(group = 1), shape = 16, size = 3, color = "red") +
  labs(title = "Motor Vehicle Theft Over Years",
       x = "Year",
       y = "Count of Incidents")
```
Motor vehicle theft shows a decline in 2019 and then increases during the pandemic and reaches  a peak level in 2023.

```{r}
ggplot(month_count, aes(x = MONTH, y = n, color = YEAR, group = YEAR)) +
  geom_line(size = 0.75) +
  geom_point(size = 2) +
  labs(title = "Monthly Total Crime Counts Over Years",
       x = "Month",
       y = "Total Crime",
       color = "Year") +
  theme_minimal() +
  theme(text = element_text(size = 12)) +
  scale_x_continuous(breaks = seq(1, 12, 1))  # Set breaks for every month
```
First four months of the year show a lower frequency of crimes (lowest in February and March) compared to the rest of the year. 2019 has the highest frequency followed by 2018 and 2023. All the years show a similar trend over the months but they differ in total number of crimes.

```{r}
ward_count <- crime %>%
  group_by(WARD) %>%
  count(OFFENSE) %>%
  summarize(sum(n))
ward_count
```

```{r}
ward_byear <- crime %>%
  group_by(YEAR, MONTH, WARD) %>%
  count(OFFENSE) %>%
  summarize(sum(n))
ward_byear
```

```{r}
ggplot(data = ward_count, aes(x = WARD, y = `sum(n)`)) +
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  geom_text(aes(label = scales::comma(`sum(n)`)), vjust = -0.5, color = "black", size = 4) +  # Add labels on top of the bars
  labs(title = "Total Crime by Ward",
       x = "Ward",
       y = "Total Crime") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1, 8, 1))
```
Ward 3 and 4 show the lowest number of crimes due to the high socioeconomic status and presence of wealthy residents. Ward 2, 5 and 1 have the highest number of crimes. Ward 2 has the highest median family income and it is not surprising that the highest number of crimes were committed there. Ward 7 and 8 has the lowest median family income and it also indicates the presence of violent crimes there. 

```{r}
methods <- crime %>%
  group_by(METHOD) %>%
  count()
methods
```

```{r}
ggplot(data = methods, aes(x = METHOD, y = `n`)) +
  geom_bar(stat = "identity", fill = "lightpink", color = "black") +
  labs(title = "Frequency of Methods used in Crimes",
       x = "Method",
       y = "Frequency") +
  theme_minimal()
```
Other methods were used more frequently than gun and knife. 

We add the CPI and Unemployment rate and we get from data.bls.gov.
```{r}
data_CPI <- read_delim("/Users/pikaqiu/Desktop/AU /class /R studio/CPI2018-2023.csv", delim = ",", col_types = cols(
  MONTH_YR = col_character(),
  Crime_number = col_double(),
  Unemploynemnt_rate = col_double(),
  CPI = col_double()
  ))
data_CPI
data_CPI <- na.omit(data_CPI)
data_CPI<- data_CPI %>%
  mutate(MONTH_YR = format(as.Date(paste0(MONTH_YR, "-01"), format = "%Y-%m-%d"), "%Y-%m"),
         Crime_number = as.integer(Crime_number))
data_CPI1 <- data_CPI |>
  filter(MONTH_YR < 2020)
data_CPI1
data_CPI2 <- data_CPI |>
  filter(MONTH_YR  > 2022)
data_CPI2

data_CPI3 <- data_CPI |>
  filter(MONTH_YR  < 2022)|>
  filter(MONTH_YR  > 2020)
data_CPI3
```

Graph for CPI and unemployment rate and Crime number
```{r}
data_CPI$MONTH_YR <- as.Date(paste0(data_CPI$MONTH_YR, "-01"))
data_CPI$Year <- format(data_CPI$MONTH_YR, "%Y")

# Create a line plot using ggplot2
ggplot(data_CPI, aes(x = MONTH_YR, y = CPI, color = Year, group = Year)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(title = "CPI Over Time",
       x = "Month",
       y = "CPI") +
  theme_minimal()

ggplot(data_CPI, aes(x = MONTH_YR, y = Unemploynemnt_rate, color = Year, group = Year)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(title = "Unemployment Rate Over Time",
       x = "Month",
       y = "Unemploynemnt_rate") +
  theme_minimal()

ggplot(data_CPI, aes(x = MONTH_YR, y = Crime_number, color = Year, group = Year)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(title = "Crime_number Over Time",
       x = "Month",
       y = "Crime_number") +
  theme_minimal()
```

Regression(linear Regression)

```{r}
# all 2018-2023
crime.reg1 <- lm(Crime_number ~  Unemploynemnt_rate + CPI, data= data_CPI)
summary(crime.reg1)
vif(crime.reg1)
qqnorm(crime.reg1$resi)
qqline(crime.reg1$resi)
```
Model for 2018-2023:
Unemployment Rate (-117): For every one-unit increase in the unemployment rate, we expect a decrease of 117 in the monthly crime frequency, holding other variables constant.

CPI (-82): For every one-unit decrease in the Consumer Price Index (CPI), we expect a decrease of 82 in the monthly crime frequency, holding other variables constant.

Both the unemployment rate and CPI have negative coefficients, indicating that higher unemployment and lower CPI are associated with lower monthly crime frequencies.

The model for 2018-2023 has follow the normal distirbution.


```{r}
# only 2018 and 2019
crime.reg <- lm(Crime_number ~  Unemploynemnt_rate + CPI, data= data_CPI1)
summary(crime.reg)
vif(crime.reg )
```
Model for 2018-2019 (Before the Covid-19 Pandemic):

Unemployment Rate (-498): For every one-unit increase in the unemployment rate, we expect a decrease of 498 in the monthly crime frequency, holding other variables constant.

CPI (1121): For every one-unit increase in the Consumer Price Index (CPI), we expect an increase of 1121 in the monthly crime frequency, holding other variables constant.

The negative unemployment coefficient suggests a decrease in crime with rising unemployment before the pandemic.The positive CPI coefficient implies an increase in crime with higher CPI during the pre-pandemic period.


```{r}
# only 2022-2023
crime.reg2 <- lm(Crime_number ~  Unemploynemnt_rate + CPI, data= data_CPI2)
summary(crime.reg2)
vif(crime.reg2)
```
Model for 2022-2023 (After the Covid-19 Pandemic):

Unemployment Rate (-500): For every one-unit increase in the unemployment rate, we expect a decrease of 500 in the monthly crime frequency, holding other variables constant.

CPI (-427): For every one-unit increase in the Consumer Price Index (CPI), we expect a decrease of 427 in the monthly crime frequency, holding other variables constant.

In the 2022-2023 model (post-pandemic), similar to the 2018-2023 model, both the unemployment rate and CPI have negative coefficients, indicating that higher unemployment and lower CPI are associated with lower monthly crime frequencies.


Conclusion:

The study concludes that the COVID-19 pandemic had a profound impact on socio-economic conditions and crime patterns in Washington D.C.. In 2020, despite the unemployment rate being at an extremely high level ~14%, there was a sharp decrease in the crime rates in Washington D.C. This was an effect of strict lockdown measures put in place by the government.When looking at crime offenses throughout the years, 2018 and 2019 overall have a higher frequency in overall crime. Since the pandemic, a contrasting pattern emerged in motor vehicle thefts, particularly for Hyundai and Kia cars, which more than doubled by 2023 from 2018. This surge is partly attributed to viral TikTok videos detailing how these vehicles can be easily stolen using simple tools like a USB. The trend was predominantly driven by younger demographics, with many inner-city youths showing a worrying disregard for the legal consequences. Homicide as well has had a sharp increase from 2018 to 2023. The rise in the cost of living, alongside the gentrification process, contributed to changing social dynamics and potential increases in certain types of crime. The findings underscore the complex interplay between economic conditions, technological advancements, and urban crime trends during and after the pandemic. 


